---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# peeler

<!-- badges: start -->
[![R-CMD-check](https://github.com/galenholt/peeler/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/galenholt/peeler/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Peeler implements the bvstep algorithm from Clarke and Warwick 1998 and uses it to 'peel' a dataset to find structural redundancy. It also provides a way to randomly start bvstep many times to assess consistency and avoid local optima.

## Installation

You can install the development version of peeler from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("galenholt/peeler")
```

## Example



```{r example}
library(peeler)
library(vegan)
data(varespec)
```

You can use `bvstep` alone

```{r}
bvout <- bvstep(ref_mat = varespec, comp_mat = varespec,
                   ref_dist = 'bray', comp_dist = 'bray',
                   rand_start = TRUE, nrand = 5)
bvout
```

The `bv_multi` function runs bvstep for a number of random starts to avoid local optima, here with 5 species to start, iterated 10 times.

```{r}
bv_m <- bv_multi(ref_mat = varespec, comp_mat = varespec,
                   ref_dist = 'bray', comp_dist = 'bray',
                   rand_start = TRUE, nrand = 5, num_restarts = 10)
bv_m
```

And finally, `peel` runs `bv_multi` iteratively, removing the best set each time.
```{r}
peels <- peel(ref_mat = varespec,
                comp_mat = varespec,
                nrand = 6,
                num_restarts = 10,
                corr_method = 'spearman')
peels
```

There are a number of user-defineable options in each of those functions, see their documentation.

THere are also two potentially useful helper functions, `extract_final`, which gets the last step in `bvstep` output, and `extract_names`. The `bvstep` output has names in a single string with comma-separated species names, and we often want them as a character vector. The `extract_names` function parses this.

```{r}
best_bv <- extract_names(bvout, step = 'last')
best_bv
```



